\vspace{30mm} {\Large\bf RISC-V Duo-PLIC for the Advancced Interrupt Architecture} \\ {\large Document Version \DuoPLICRev}

### 2023-03-11 doc/src/MSLevel.tex
Allocate local interrupts for RAS events
Signed-off-by: John Hauser <31252952+jhauser-us@users.noreply.github.com>

Table~\ref{tab:majorIntrs} lists all the major interrupts
currently defined for {\RISCV} harts.
At this time, the only standard major interrupts are those
specified by the {\RISCV} Privileged Architecture.

\begin{table*}[h!]
\begin{center}
\begin{tabular}{|c|l|l|}
\hline
Default        &                         & \\
priority order & Major interrupt numbers & \multicolumn{1}{c|}{Description} \\
\hline
\hline
Highest & 11, 3, 7   & Machine interrupts:  external, software, timer \\
        & \ 9, 1, 5  & Supervisor interrupts:  external, software, timer \\
        & 12         & Supervisor guest external interrupt \\
        & 10, 2, 6   & VS interrupts:  external, software, timer \\
Lowest  & 13         & Counter overflow interrupt \\
\hline
\end{tabular}
\end{center}
\caption{%
The standard major interrupt codes, listed in default priority order.
Currently, all standard codes are defined by
the {\RISCV} Privileged Architecture.%
}
\label{tab:majorIntrs}
\end{table*}

\begin{table*}[h!]
\begin{center}
\begin{tabular}{|c|l|}
\hline
                          & \\
Major interrupt numbers   & \hspace{10mm}Category \\
\hline
\hline
\begin{tabular}{@{}c@{}}
  \ 0--12 \\
  13--15 \\
\end{tabular} &
    \begin{tabular}{@{}l@{}}
      Not local interrupts \\
      Local interrupts \\
    \end{tabular}
      $\left\}\begin{tabular}{l@{}}
        Assigned by the \\
        \quad Privileged Architecture \\
      \end{tabular}\right.$ \\
\hline
16--23                    & Local interrupts\\
24--31                    & \em Designated for custom use \\
32--47                    & Local interrupts \\
\mbox{\ $\geq \mbox{48}$} & \em Designated for custom use \\
\hline
\end{tabular}
\end{center}
\caption{Categorization of current and future major interrupts.}
\label{tab:majorIntrCategories}
\end{table*}

Of the major interrupts controlled by the Privileged Architecture
(numbers 0--15), this Advanced Interrupt Architecture (AIA) categorizes the
counter overflow interrupt (code~13) as a \emph{local interrupt}.
It is assumed furthermore that any future definitions for
reserved interrupt numbers 14 and 15 will also be local interrupts.
The AIA additionally reserves
major interrupt numbers 16--23 and 32--47 for
standard local interrupts that other {\RISCV} extensions may define.
The remaining major interrupts allocated to the Privileged
Architecture, numbers 0--12, are categorized as not local interrupts.
Taken altogether, Table~\ref{tab:majorIntrCategories}
summarizes the AIA's categorization
of all major interrupt identities.

@@

\begin{commentary}
In addition to the existing major interrupts of
Table~\ref{tab:majorIntrs}, the following
local interrupts are tentatively proposed,
listed in order of decreasing default priority:\nopagebreak
\begin{displayLinesTable}[l@{\quad}l]
23 & Bus or system error \\
45 & Per-core high-power or over-temperature event \\
43 & High-priority RAS event \\
\noalign{\smallskip}
35 & Low-priority RAS event \\
17 & Debug/trace interrupt \\
\end{displayLinesTable}
\noindent
These local interrupts are expected to be
specified by other {\RISCV} extensions.
Be aware, this list is not final and may change
as the relevant extensions are developed and ratified.
\emph{RAS} is an abbreviation for \emph{Reliability, Availability, and
Serviceability}.
\end{commentary}

>>>>

Table~\ref{tab:majorIntrs} lists all the major interrupts
currently defined for {\RISCV} harts
that conform to this Advanced Interrupt Architecture (AIA).
Besides the major interrupts specified by the {\RISCV} Privileged
Architecture, the AIA adds interrupt numbers 35 and 43
as local interrupts for low- and high-priority \emph{RAS events}.

\begin{table*}[t]
\begin{center}
\begin{tabular}{|c|l|l|}
\hline
Default        &                         & \\
priority order & Major interrupt numbers & \multicolumn{1}{c|}{Description} \\
\hline
\hline
Highest & 43         & Local interrupt: high-priority RAS event \\
        \cline{2-3}
        & 11, 3, 7   & Machine interrupts:  external, software, timer \\
        & \ 9, 1, 5  & Supervisor interrupts:  external, software, timer \\
        & 12         & Supervisor guest external interrupt \\
        & 10, 2, 6   & VS interrupts:  external, software, timer \\
        & 13         & Local interrupt: counter overflow \\
        \cline{2-3}
Lowest  & 35         & Local interrupt: low-priority RAS event \\
\hline
\end{tabular}
\end{center}
\caption{%
The standard major interrupt codes, listed in default priority order.%
}
\label{tab:majorIntrs}
\end{table*}

\begin{table*}[t]
\begin{center}
\begin{tabular}{|c|l|}
\hline
                          & \\
Major interrupt numbers   & \hspace{10mm}Category \\
\hline
\hline
\begin{tabular}{@{}c@{}}
  \ 0--12 \\
  13--15 \\
\end{tabular} &
    \begin{tabular}{@{}l@{}}
      Not local interrupts \\
      Local interrupts \\
    \end{tabular}
      $\left\}\begin{tabular}{l@{}}
        Assigned by the \\
        \quad Privileged Architecture \\
      \end{tabular}\right.$ \\
\hline
16--23                    & Local interrupts\\
24--31                    & \em Designated for custom use \\
32--47                    & Local interrupts \\
\mbox{\ $\geq \mbox{48}$} & \em Designated for custom use \\
\hline
\end{tabular}
\end{center}
\caption{Categorization of current and future major interrupts.}
\label{tab:majorIntrCategories}
\end{table*}

Of the major interrupts controlled by the Privileged Architecture
(numbers 0--15), the AIA categorizes the
counter overflow interrupt (code~13) as a \emph{local interrupt}.
It is assumed furthermore that any future definitions for
reserved interrupt numbers 14 and 15 will also be local interrupts.
Besides the two RAS interrupts, the AIA additionally reserves
major interrupt numbers in the ranges 16--23 and 32--47 for
standard local interrupts that other {\RISCV} extensions may define.
The remaining major interrupts allocated to the Privileged
Architecture, numbers 0--12, are categorized as not local interrupts.
Taken altogether, Table~\ref{tab:majorIntrCategories}
summarizes the AIA's categorization
of all major interrupt identities.

\emph{RAS} is an abbreviation for
\emph{Reliability, Availability, and Serviceability}.
Typically a RAS event corresponds to the detection
of corrupted data (e.g.\ as a result of a soft or hard error)
and/or the use of such data.
The high-priority RAS event local interrupt may, for example,
signal an occurrence of an urgent uncorrected error
that needs action from a RAS error handler to
contain the error and, if possible, to recover from it.
The low-priority RAS event local interrupt may, for example,
be triggered by non-urgent deferred or corrected errors.

The AIA does not itself require that detected RAS events trigger
one of the two local interrupts defined for this purpose.
Systems are free to report any or all RAS events
another way, such as by external interrupts routed
through an APLIC or IMSIC, or by custom interrupts.
In all likelihood, the method for reporting a particular RAS event
will depend on where in the system the event is detected.
The AIA defines local interrupt numbers for RAS events so systems
have a standard way to report such events when detected locally at
a hart, without depending solely on external or custom interrupts.

\begin{commentary}
As always, platform standards may further constrain how
a system reports events, whether RAS events or other.
\end{commentary}

@@

\begin{commentary}
In addition to the existing major interrupts of
Table~\ref{tab:majorIntrs}, the following
local interrupts are tentatively proposed,
listed in order of decreasing default priority:\nopagebreak
\begin{displayLinesTable}[l@{\quad}l]
23 & Bus or system error \\
45 & Per-core high-power or over-temperature event \\
17 & Debug/trace interrupt \\
\end{displayLinesTable}
\noindent
These local interrupts are expected to be
specified by other {\RISCV} extensions.
Be aware, this list is not final and may change
as the relevant extensions are developed and ratified.
\end{commentary}
